Data Specialization for a Staged Lambda Calculus

### 2015-02-26 paper/splitting.tex
Pass through 4.1
Decided to remove "holes" and just keep text saying they were removed.

  Masking converts a partial value and associated residual table into a
masked value $\mval i q$, where $i$ is a monostage value encoding all
first stage components of the partial value, and $q$ is a monostage
term encoding all of the second-stage computations of the partial
value.  Remember that masking will not play a role in the final
splitting algorithm, but instead exists to specify the contract to
which splitting at \bbonem\ must adhere.  Accordingly, masking should
be thought of as a definition, and as such it is in some parts
arbitrary.

So to motivate the definition we chose, consider the following residual table and partial value,
colored as before:
\begin{lstlisting}
2`[yhat |-> 1+2]`1` (next{`2`yhat`1`}, injL (grnd{7}, next{`2`yhat`1`}))`
\end{lstlisting}
To reduce this term to just its first-stage information ($i$ in the output), we start by redacting all of the second stage (blue) parts,
along with the surrounding \texttt{next} annotations:
\begin{lstlisting}
4`##########`1` (`4`######`1`, injL (grnd{7}, `4`######`1`))`
\end{lstlisting}
The holes inside the term needs to be filled with something to convey the lack of information, so we use a unit values:
\begin{lstlisting}
1`((), injL (grnd{7},()))`
\end{lstlisting}
Finally, we drop \texttt{grnd} annotations:
\begin{lstlisting}
1`((), injL (7,()))`
\end{lstlisting}
To reduce the original partial value to just its second-stage information ($q$ in the output), 
we start by redacting all \texttt{grnd} blocks, \texttt{next} annotations, and injection tags
(essentially, the tuple structure is kept, but all other structure is discarded):
\begin{lstlisting}
2`[yhat |-> 1+2]`1` (`4`####`2`yhat`4`#`1`, `4`####`1` (`4`####`1`,`4`####`2`yhat`4`#`1`))`
\end{lstlisting}
Where the \texttt{grnd} blocks left a hole, we put in a unit values:
\begin{lstlisting}
2`[yhat |-> 1+2] (yhat, ((),yhat))`
\end{lstlisting}
Finally, we reify the residual table into \texttt{let} bindings:
\begin{lstlisting}
2`let yhat = 1+2 in (yhat, ((),yhat))`
\end{lstlisting}

The formal implementation of the masking relation is given in \ref{fig:valMask}.
Note that since lambdas may represent multistage computations, 
masking splits the body of lambdas as general world \bbonem\ terms (as described in \ref{sec:split-one}), 
and packages the resulting terms as functions in both the first and second stage.
% Masking operates by first inducting on the entries of the residual table.  
% Being purely second-stage content, these are reified into let statements at the top of the resumer.
% Once the table is empty, masking inducts on value itself.

>>>>

Masking converts a residual table and associated partial value into a
masked value $\mval i q$.
%that explicitly separates the first stage components of 
%result from the residual.
%where $i$ is a monostage value encoding all
%first stage components of the partial value, and $q$ is a monostage
%term encoding all of the second-stage computations of the partial
%value.
Masking is not a transformation performed as part of the splitting
algorithm.  Instead, masking defines an intuitive notion of the first
stage and second stage ``components'' of residual tables and partial
values, and it is used to define the requirements of what output
splitting of \bbonem\ terms must produce.  A precise definition of the
masking relation is given in \ref{fig:valMask}, but to gain intuition
about the intent of masking operations, consider the following
residual table and partial value, colored as before:
\begin{lstlisting}
2`[yhat |-> 1+2]`1` (next{`2`yhat`1`}, injL (grnd{7}, next{`2`yhat`1`}))`
\end{lstlisting}
To construct a value representing its first-stage components ($i$ in
the resulting masked value), masking first redacts all second stage
(blue) parts, along with the surrounding \texttt{next}
annotations. (This redaction removes the entire residual table since
it only represents stage two computations.) The resulting ``holes'' in
the term are replaced with unit values.
%\begin{lstlisting}
%4`##########`1` (`4`######`1`, injL (grnd{7}, `4`######`1`))`
%\end{lstlisting}
%The holes inside the term needs to be filled with something to convey the lack of information, so we use a unit values:
\begin{lstlisting}
1`((), injL (grnd{7},()))`
\end{lstlisting}
Finally, masking dropping \texttt{grnd} annotations yielding a value.
\begin{lstlisting}
1`((), injL (7,()))`
\end{lstlisting}
To construct a residual corresponding to the stage two computations
represented by the residual table and partial value ($q$ in resulting
the masked value), masking first redacts all \texttt{grnd} blocks
(replaced with unit), \texttt{next} annotations, and injection tags
(essentially, the tuple structure is kept, but all other structure is
discarded):
%\begin{lstlisting}
%2`[yhat |-> 1+2]`1` (`4`####`2`yhat`4`#`1`, `4`####`1` (`4`####`1`,`4`####`2`yhat`4`#`1`))`
%\end{lstlisting}
%Where the \texttt{grnd} blocks left a hole, we put in a unit values:
\begin{lstlisting}
2`[yhat |-> 1+2] (yhat, ((),yhat))`
\end{lstlisting}
Then masking reifies the residual table into \texttt{let} bindings,
yielding a residual:
\begin{lstlisting}
2`let yhat = 1+2 in (yhat, ((),yhat))`
\end{lstlisting}

Note in \ref{fig:valMask} that since lambdas may represent multistage
computations, masking splits the body of lambdas as general world
\bbonem\ terms (as described in \ref{sec:split-one}), and packages the
resulting terms as functions in both the first and second stage parts
of the masked value.
% Masking operates by first inducting on the entries of the residual table.  
% Being purely second-stage content, these are reified into let statements at the top of the resumer.
% Once the table is empty, masking inducts on value itself.
